Night and day: What are the nonlinear associations between urban characteristics and land surface temperature?

### 2019-04-30 report/article.tex
Updates from Overleaf

\author[2]{B. Zaitchik}
\author[1]{S. Guikema}

@@

\begin{abstract}
Although heat waves and the urban heat island are nocturnal phenomena, the drivers of land surface temperature during the night remain poorly understood.
Understanding these drivers is necessary for reducing land surface temperature and mitigating heat waves, the deadliest natural hazard, which are expected to increase in frequency and severity.
% However, only recently has nocturnal imagery become available from LandSat allowing nighttime land surface temperature to be analyzed.
However, the independent effects and relative importance of potential drivers remain unclear from existing studies.
We seek to answer the question: what are the effects and importance of drivers of land surface temperature during both day and night?
To do this, we analyze the urban land surface temperature in four cities across the United States.
In our analysis, we include factors related to vegetation, water, the built-environment, and topography and control them using nonlinear statistical methods which allow for their independent effects to be assessed.
The effects from the diurnal and nocturnal analysis are compared to determine if previously reported relationships hold between cities and during both the day and night.
Understanding the relationships influencing both nocturnal and diurnal land surface temperature will improve climate adaptation planning and heat wave mitigation.
\end{abstract}

\begin{keyword}
Urban climate; LandSat; Land surface temperature; Urban canyon; Machine learning
\end{keyword}

@@

\section{Introduction}
In a warming world, understanding the drivers of high land surface temperature (LST) and the urban heat island (UHI) will aid in adapting cities in mitigating urban heat for the health and well being of their communities.
And mitigate they must, the 1995 Chicago heat wave, which killed more than 700 people\footnote{The 2003 European heatwave killed 70,000 \cite{Robine2008-ky} and the 2015 European heat wave increased mortality up to 30\% \cite{Vicedo-Cabrera2016-si}} is expected to become an annual occurrence by 2080 \cite{klinenberg2015heat}. 
Heat waves' affect on people is exacerbated by the urban heat island (UHI) \cite{Wicki2017-fv, Echevarria_Icaza2016-fr}.
The UHI results from heat captured during the day being released during the night, thus increasing nocturnal temperatures \cite{Oke2002-ta, Landsberg1981-mq, Rotach2005-yu}.
This increase in nocturnal temperature reduces people's ability to cool off during the night and drives an increase in mortality \cite{Echevarria_Icaza2016-fr, Murage2017-wj}.

Urban heat is studied in two ways.
The first uses weather station data to analyze air temperature \cite{Scott2016-lc}(etc.) while the other uses remote sensing to examine the land surface temperature (LST) \cite{Imhoff2010-lf, Peng2012-iy, Peng2018-cp, Zhou2014-wc}.
Although there are differences between land temperature and air temperature \cite{Good2016-yk}, the extensive spatial coverage of LST data is a major advantage and enables comparative studies \cite{Hung2006-qy}.
For the sake of comparative analysis, we use LST in this study.

Most existing LST studies focus on the daytime \cite{Peng2018-cp,Chun2018-so,Wang2019-water,Zhou2018-iy}.
However, the mechanisms and urban characteristics driving LST allegedly differs between night and day \cite{Hung2006-qy, Chun2017-mm, Nichol2005-mm, Wicki2017-fv, Echevarria_Icaza2016-fr,Sobstyl2018-wt, Peng2012-iy, Zhou2014-wc, Zhao2017-cc}. 
Given that UHI is a nocturnal phenomena and the lack of studies focused on nocturnal LST, we are yet to fully understand what drives LST during the night.

The existing studies on nocturnal LST are yet to provide a clear understanding of the the influence and relative importance of urban characteristics associated with land surface temperature \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}. 
These studies can be enhanced in three ways: 1) studying multiple regions, 2) considering 2D and 3D urban characteristics, 3) improving their statistical techniques.
The data availability has limited previous nocturnal studies, meaning that many rely on MODIS images with a 1km resolution \cite{Zhou2014-wc, Echevarria_Icaza2016-fr,Wang2019-tree,Peng2012-iy}.
However, LandSat8 (L8) night scenes have recently become available. 
Due to the wide spatial availability of LandSat imagery, comparative studies between cities, necessary to understand the generalizability of findings \cite{Peng2012-iy, Hung2006-qy}, can now be conducted for day and night.

Existing studies have also been criticized for the explanatory variables they've used. [who's criticised them?]
There is ongoing disagreement regarding the importance of 3D (e.g. building height) vs 2D (such as albedo) variables.  
Competing studies suggest that 3D factors are not important \cite{Berger2017-lx}, while others find that ignoring 3D incorrectly conflates the effect of different 2D variables \cite{Chun2017-mm}.
Beyond the 2D or 3D question, five categories of important variables have been identified: Green space, Water, Landscape, Albedo, and Socio-economic \cite{Peng2018-cp}. 
However, many studies do not capture all categories and, worse, many analyze only the univariate effect of variables \cite{Zhao2017-cc, Merbitz2012-xz, Unger2004-ry} (see \cite{Peng2018-cp, Chun2017-mm} for further discussion). 
Considering only single variables limits the understanding of variable effect given the interdependencies between drivers.

The third potential enhancement is in their statistical approaches.
* multicollinearity problems and multiple statistical techniques might be need to enhance reliability and interpretation \cite{Zhou2014-wc}
The current approaches used in the studies limit their ability to explain the interdependencies between variables. 
Almost all studies continue to use linear techniques (e.g. \cite{Li2017-yl, Peng2012-iy, Wicki2017-fv,Zhou2014-wc,Peng2018-cp,Echevarria_Icaza2016-fr,Chun2017-mm,Chun2018-so,Wang2019-tree,Wang2019-water}, making quantifying the independent effects of factors on LST and their relative influence difficult \cite{Peng2018-cp, Zhou2014-wc}.
These studies also only assess their accuracy using in-sample validation techniques.
In-sample validation means that model accuracy is assessed with the same data used in the training, rather than unseen data; the result is potential overestimation of explained variance (e.g. R$^2$ or other validation metric).

Therefore, in this study we determine the influence and relative importance of urban characteristics associated with LST, during both the night and day. 
To achieve this we conduct a comparative study of four cities in the United States using non-linear statistical techniques which capture the interdependencies and relative importance between the urban characteristics.
The objective is to assess whether the characteristics thought to influence high temperatures during the day are also influencing temperatures at night.
This has implications for mitigating the severity of future heat waves.

@@

\subsection{Land surface temperature}
We calculate land surface temperature using Landsat 8 (Land Processes Distributed Active Archive Center product) imagery as well as land cover and an air temperature observation (data sources provided in \ref{tab:data}. Band-10 digital-number data is converted to top-of-atmosphere radiance \cite{Jimenez-Munoz2003-wc}. We correct for emissivity using land cover data \cite{Alipour2003-gb}, and then calculate the at-satellite brightness temperature \cite{Jimenez-Munoz2003-wc}. Finally an atmospheric correction is made as per the monowindow algorithm \cite{Qin2001-jn} using the maximum observed temperature of the day from a nearby weather station. This follows the process is described in \cite{Scott2016-lc} and is explicit in our open source code on Github\footnote{URL redacted for review}.

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=\linewidth]{fig/report/scatter_500.png}
    \caption{The land surface temperature in $^o$C of the cities studied at a 500m square resolution}
    \label{fig:scatter_lst}
    \end{center}
\end{figure}

@@

\textit{Albedo}. Albedo is a measure of the reflectivity or brightness of a surface. It is normalized between 0-1 where the darker surfaces are lower values. Albedo is calculated using the LandSat8 images using the algorithm described in \cite{Smith2010-nw, Liang2001-jd}.

@@

\textit{Building height}. Building height is estimated using the building footprints and available lidar data following \cite{Chun2017-mm}. The mean lidar elevation within each building's footprint is subtracted by the mean digital elevation (topography) within the footprint to estimate the building's height.

\textit{Elevation}. 1/3 arc-second (~10m) bare-earth elevation (topography) data is available courtesy the U.S. Geological Survey. The mean, max, and min of elevation points within the grid cell is calculated.

@@

\textit{Impervious surface percentage}. This is provided in the National Land Cover Database from the Multi-Resolution Land Charasteristic Consortium \cite{Xian2011-aa}. To generate the impervious surface area (ISA) LandSat data, NLCD land cover, and nighttime light imagery is used. The stable nighttime light intensity is only used to estimate the boundary of of urban areas. The Landsat images were converted to top-of-atmosphere reflectance. The data is provided at a 30m resolution and we calculate the mean, max, and min for each grid cell.

**** show scatter against tree canopy - cannot include both ****

\textit{Land cover}. Classification of land cover is described in \cite{Homer2015-ce}. However, as it is used in the calculation of emissivity when calculating the LST it cannot be used as an independent variable. The only landcover that we use is category 11, water. Cells with more than 20\% area classified as water are excluded from the data set.

@@

\textit{Population density}. Density allegedly increases the LST \cite{Li2017-yl}, although this result may be the result of confounding with other factors. Population density calculated from the U.S. census at the block level. The most recent census was 2010, so that data is used as an estimator of where people reside in the evening. This data is also at a lower resolution that the grid cells used, so the grid cell assumes the density of the block that it's centroid is contained within.

\textit{Sky view factor}. Urban canyons have been found to have an effect on UHI because they prevent air circulation \cite{Landsberg1981-mq, Chun2017-mm} and are used to indicate radiation flux within complex environments \cite{Matzarakis2007-xy}. We calculate SVF following using xxx and following the parameters used by \cite{Chun2017-mm}: the number of search directions, $\phi=10^o$; and the radius of the reference circle, $R=300m$. We use a spatial resolution of x-m.
% * Yuan 2011, Unger 2004.

\textit{Tree canopy cover}. The percent tree canopy cover is calculated using National Agriculture Imagery Program (NAIP) aerial imagery, Landsat 5 imagery, elevation, and existing NLCD data \cite{Coulston2012-uu, Homer2015-ce}. The data provided is at an xx resolution. The six reflective bands from Landsat 5 are used to calculate top-of-atmosphere reflectance \cite{Coulston2012-uu} so the data does not contain information used in the LST calculation (which requires radiance). 
% * Rogan 2013, Elmes 2017

\subsection{Data preparation and robustness}

To analyze the data, it needs to transformed into a spatially consistent set.
The raw data is a variety of spatial data types from area level (the population density data), to geostatistical raster (e.g. the land surface temperature).
We choose to resample the data into a grid of square cells.
We conduct the resampling twice to produce a grid of 500m and 100 meter cells. 
Having two grid resolutions allowed us to ensure our conclusions are robust.
For each of the cells the the mean, maximum, and minimum of all variables were calculated. 
To further account for spatial effects of each variable, the mean of the surrounding cells (including diagonal) was calculated and the resulting spatial lag variable was included as an additional independent variable. 
The result is a data set which with we can train a variety of statistical models.

Various statistical models (\S \ref{ss:models}) were trained and then, crucially, validated on unseen data using a technique known as holdout cross-validation. 
This approach partitions 80\% of the data into a training set and the remaining 20\% into the testing set.
The model resulting from the training, is then tested on this unseen test set, to get an estimate of the model accuracy.
This is repeated 100 times and the distribution of the accuracy metric (here we use the mean absolute error (MAE) and variance explained (R$^2$)).
Holdout cross-validation of this type is crucial for statistical analysis to ensure that models are not overfit to data \cite{Geron2017-ek}.
Overfitting occurs when a model fits to the randomness in a dataset, causing it to not be suitable for generalizations. 
When models' accuracy metrics are reported based on in-sample data (the same data it was trained on), the accuracy metrics are high.
This is essential for statistical models and overlooked by the majority of existing studies (e.g. \cite{Zhou2014-wc, Peng2018-cp, Chun2017-mm, Chun2018-so,Wang2019-tree}).
Cross-validation helps to avoid this by evaluating the model on unseen data.

@@

\textit{Random Forest}. A random forest is an ensemble model of \textit{Regression Trees}. A regression tree partitions the data based on thresholds for the covariates \cite{Breiman1984-hw}. It continues to divide the data on different covariates until the partitions are as similar as possible. 
The result is a tree-like structure, after which the model is named.
In similar naming fashion, a random forest is a collection of regression trees.
Each tree is \textit{grown} from a random subset of the data and the prediction from the random forest model is the average of the prediction from all of the trees \cite{Breiman2001-rt}.

@@

\section{Results and Discussion}

During the day, Increased impervious surfaces and decreased vegetation increase sensible heat flux and lower latent heat flux \cite{Voogt2003-mm, Peng2012-iy, Zhou2014-wc}

Heat storage during the day, increases temperatures during the night \cite{Voogt2003-mm, Zhou2014-wc}

Day and night mechanisms that drive temperature differ, likely because latent and sensible heat dominant during the day, but ground heat flux dominates at night \cite{Zhou2014-wc, Voogt2003-mm}.

Vegetation and pervious surfaces caused the greatest reduction in land surface temperature, with approximately 10oC change during the day (Figure \ref{fig:pdp_2dday_500}b). 
This result supports existing findings \cite{Zhou2014-wc, Chun2018-so}.
it is likely due to vegetation increasing the latent heat fluxes due to transpiration \cite{Zhou2014-wc}, which explains why it has a greater effect during the day. 
vegetation does not transpire during the night however, so the reduction in temperature due to increased tree cnaopy cover, is likely due to a reduction in impervious surfaces that store heat \cite{Voogt2003-mm}. 
Not all types of pervious surfaces are as good as others \cite{Gober2009-im}. Care must be taken when irrigating \cite{Gober2009-im}.
urban design needs to consider interaction between surface materials \cite{Guhathakurta2010-ib}.
- vegetation important during the day \cite{Chun2017-mm, Peng2018-cp}
- vegetation is important \cite{Wang2019-tree}

water
- showed no effect in the 100m data, because so few data observations had water in them.
- dramatic reductions in temperature during both night and day (y $^o$C).
- this contrasts with findings and rationale that water releases heat at night \cite{Chun2017-mm} - although that logic applies to air temperature, not sure how strong water would be at increasing temperature of nearby land surface
- supports strong negative findings \cite{Wang2019-water}

Urbanization leads to heat storage in roads and buildings \cite{Zhou2014-wc, Voogt2003-mm}. 
the results were surpringly inconsistent, regarding albedo - however, this is confounded as vegetation and highly impervious surfaces are both dark (low albedo).
considering 2d partial dependence, we see that albedo has little affect at night it does show that the highest temperatures are the most built up and the least reflective (Figure \ref{fig:pdp_2dnight_500}a).
during the day, tincreasing the whiteness, decreases the temperature. the higest temperatures are observed when there is significant building area and low albedo (Figure \ref{fig:pdp_2dday_500}a) and high impervious area and low albedo (Figure \ref{fig:pdp_2dday_500}b).
- albedo, then ndbi to be most important, only consider day \cite{Peng2018-cp}
- role of impervious surface \cite{Lu2006-ch}

The canyon effect is also attributed to cause warmer temperatures. \cite{Chun2017-mm,Oke1988-re}
while there was high uncertainty in the models, it appears that during the night, the temperatures decrease as the sky view factor increased (Figure \ref{fig:pdp_night_500}d).
This is potentially due to heat being stored within the canyons (areas with low, sky view factors).
Conversely, during the day, areas with high sky view factor had increased temperature (Figure \ref{fig:pdp_day_500}d).
This is may be a result of shade from the buildings \cite{Yu2019-ps}.
Urban canyon was found to be important at night \cite{Chun2017-mm}

The building area had no affect during the night (Figure \ref{fig:pdp_night_500}) and increased temperature during the day (Figure \ref{fig:pdp_day_500}).
The population density had no discernible affect during night or day.

Supports suggestion that 2D metrics outperform 3D metrics \cite{Berger2017-lx}

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{fig/report/holdout_500.png}
    \caption[Holdout cross-validation results at 500 meter resolution]{
    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models from a 100-fold holdout cross-validation. 
    The models were trained on 80\% of the data and tested on the unseen 20\%.
    When selecting data for the training and testing sets, spatial subsets were used to account for spatial similarities. 
    OOB R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1. 
    Good models have MAE near 0.
    }
    \label{fig:holdout_500}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{fig/report/pdp_500.png}
    \caption[Partial dependence plots for LST at 500 meter resolution]{
    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each urban characteristic as the other variables are held at their average (mean) value. 
    The left hand side shows the effect each variable has on the (a) mean land surface temperature (LST) during the night, (b) maximum LST during the night, (c) mean LST during the day, (d) maximum LST during the day. 
    This is for the 500 meter resolution.
    }
    \label{fig:pdp_night_500}
\end{figure*}

>>>>

\author[2]{B. Zaitchik}
\author[1]{S. Guikema}
\author[]{A. Nisbet}

@@

\begin{abstract}
Although heat waves and the urban heat island are nocturnal phenomena, the drivers of land surface temperature during the night remain poorly understood.
Understanding these drivers is necessary for reducing land surface temperature and mitigating heat waves, the deadliest natural hazard, which are expected to increase in frequency and severity.
% However, only recently has nocturnal imagery become available from LandSat allowing nighttime land surface temperature to be analyzed.
However, the independent effects and relative importance of potential drivers remain unclear from existing studies.
We seek to answer the question: ``What is the relative importance and nonlinear affects of urban characteristics associated with land surface temperature during both the day and night?'' 
To do this, we analyze the urban land surface temperature in four cities across the United States.
In our analysis, we include variables related to vegetation, water, the built-environment, and topography. 
We control these variables using nonlinear statistical methods which allow for their independent effects to be assessed.
The effects from the daytime and nighttime analysis are compared to determine if previously reported relationships hold between cities and during both the day and night.
Understanding the relationships influencing both night and day land surface temperature will improve climate adaptation planning and heat wave mitigation.
\end{abstract}

\begin{keyword}
Urban climate; LandSat; Land surface temperature; Urban canyon; Machine learning; Convolutional Neural Network
\end{keyword}

@@

\section{Introduction}
In a warming world, understanding the drivers of high land surface temperature (LST) and the urban heat island (UHI) will aid in adapting cities in mitigating urban heat for the health and well being of their communities.
And mitigate they must; the 1995 Chicago heat wave, which killed more than 700 people,\footnote{The 2003 European heatwave killed 70,000 \cite{Robine2008-ky} and the 2015 European heat wave increased mortality up to 30\% \cite{Vicedo-Cabrera2016-si}} is expected to become an annual occurrence by 2080 \cite{klinenberg2015heat}. 
Heat waves' affect on people is exacerbated by the urban heat island (UHI) \cite{Wicki2017-fv, Echevarria_Icaza2016-fr}.
The UHI results from heat captured during the day being released during the night, thus increasing nocturnal temperatures \cite{Oke2002-ta, Landsberg1981-mq, Rotach2005-yu}.
This increase in nocturnal temperature reduces people's ability to cool off during the night and drives an increase in mortality \cite{Echevarria_Icaza2016-fr, Murage2017-wj}.

Urban heat is studied in two ways.
The first uses weather station data to analyze air temperature (e.g. \cite{Scott2016-lc}) while the other uses remote sensing to examine the land surface temperature (LST) \cite{Imhoff2010-lf, Peng2012-iy, Peng2018-cp, Zhou2014-wc, Voogt2003-mm}.
Although there are differences between LST and air temperature \cite{Good2016-yk}, the extensive spatial coverage of LST data is a major advantage and enables comparative studies \cite{Hung2006-qy}.
We analyze LST in this study.

Most existing LST studies focus on the daytime \cite{Peng2018-cp,Chun2018-so,Wang2019-water,Zhou2018-iy}.
However, the mechanisms and urban characteristics driving LST allegedly differs between night and day \cite{Hung2006-qy, Chun2017-mm, Nichol2005-mm, Wicki2017-fv, Echevarria_Icaza2016-fr,Sobstyl2018-wt, Peng2012-iy, Zhou2014-wc, Zhao2017-cc}. 
Given that UHI is a nocturnal phenomena, the lack of study on nighttime LST leaves a critical gap in our understanding. 
This can now be addressed as high resolution nighttime satellite images have become available.

Existing studies that have analyzed LST have used low-resolution imagery, making it challenging to deduce a clear understanding of the the influence and relative importance of the associated urban characteristics \cite{Chun2017-mm, Echevarria_Icaza2016-fr, Wicki2017-fv, Zhou2014-wc}. 
Additionally, these studies can be enhanced by: 1) studying multiple regions, 2) considering 2D and 3D urban characteristics, 3) improving their statistical techniques.
The data availability has limited previous nocturnal studies, meaning that many rely on MODIS images with a 1km resolution \cite{Zhou2014-wc, Echevarria_Icaza2016-fr,Wang2019-tree,Peng2012-iy}.
1km resolution makes it difficult to attribute LST to urban characteristics.
However, 30-m resolution LandSat8 (L8) night scenes have recently become available. 
Due to the wide spatial availability of L8 imagery, comparative studies between cities, necessary to understand the generalizability of findings \cite{Peng2012-iy, Hung2006-qy}, can now be conducted for day and night.

Existing studies have also been criticized for the explanatory variables they've used \cite{Chun2017-mm,Peng2018-cp}.
There is ongoing disagreement regarding the importance of 3D (e.g. building height) vs 2D (such as albedo) variables.  
Competing studies suggest that 3D factors are not important \cite{Berger2017-lx}, while others find that ignoring 3D incorrectly conflates the effect of different 2D variables \cite{Chun2017-mm}.
Beyond the 2D or 3D debate, five categories of important variables have been identified: Green space, water, landscape, albedo, and socio-economic \cite{Peng2018-cp}. 
However, many studies do not capture these categories and, worse, many analyze only the single effect of each variable \cite{Zhao2017-cc, Merbitz2012-xz, Unger2004-ry} (see \cite{Peng2018-cp, Chun2017-mm} for further discussion). 
Considering a variable in isolation, without accounting for potential conflating by other variables, limits the understanding of the interdependent effects that exist.

The third potential enhancement is in the statistical approaches.
Almost all studies use linear techniques (e.g. \cite{Li2017-yl, Peng2012-iy, Wicki2017-fv,Zhou2014-wc,Peng2018-cp,Echevarria_Icaza2016-fr,Chun2017-mm,Chun2018-so,Wang2019-tree,Wang2019-water}.
Using linear models incur a number of major challenges that limit these studies' ability to explain the interdependencies between variables. 
The first is that many of the urban characteristics exhibit high multicollinearity \cite{Zhou2014-wc}.
The second is that linear models are limited in their ability to quantify the independent effects of characteristics and their relative influence on LST \cite{Peng2018-cp, Zhou2014-wc}.
Additionally, and most importantly, these studies only assess their accuracy using in-sample validation techniques.
In-sample validation means that model accuracy is assessed with the same data used in the training, rather than unseen data. 
This risks overestimating the accuracy of their models.

The question we address is: ``What is the relative importance and nonlinear affects of urban characteristics associated with land surface temperature during both the day and night?''
To achieve this we conduct a comparative study of four cities in the United States using non-linear statistical techniques which capture the interdependencies and relative importance between the urban characteristics.
In addressing our question, we will address the, at times fundamental, limitations of existing studies.
Most importantly we look at nighttime temperature as well as daytime, we appropriately validate our statistical models, and we use tools that allow us to explore the interdependent effects of urban characteristics on urban heat.
The results have implications for mitigating the severity of future heat waves.

@@

\subsection{Land surface temperature}
We calculate land surface temperature using Landsat 8 (Land Processes Distributed Active Archive Center product) imagery, land cover data, and an air temperature observation (data sources provided in \ref{tab:data}). 
Our code is available on our Github\footnote{URL redacted for review} repository, and is as follows:
We convert Band-10 digital-number data to top-of-atmosphere radiance \cite{Jimenez-Munoz2003-wc}. 
We correct for emissivity using land cover data \cite{Alipour2003-gb}, and then calculate the at-satellite brightness temperature \cite{Jimenez-Munoz2003-wc}. 
Finally an atmospheric correction is made as per the monowindow algorithm \cite{Qin2001-jn} using the maximum observed temperature of the day from a nearby weather station. 
This follows the process is described in \cite{Scott2016-lc}.

@@

\begin{figure}
    \begin{center}
    \includegraphics[width=\linewidth]{fig/report/scatter_500.png}
    \caption{The land surface temperature in $^o$C of the cities studied at a 500m square resolution.}
    \label{fig:scatter_lst}
    \end{center}
\end{figure}

@@

\textit{Albedo}. Albedo is a measure of the reflectivity or brightness of a surface. It is normalized between 0-100 where the darker surfaces are lower values. Albedo is calculated using the LandSat8 images using the algorithm described in \cite{Smith2010-nw, Liang2001-jd}.

@@

% \textit{Building height}. Building height is estimated using the building footprints and available lidar data following \cite{Chun2017-mm}. The mean lidar elevation within each building's footprint is subtracted by the mean digital elevation (topography) within the footprint to estimate the building's height.

\textit{Elevation}. 1/3 arc-second (~10m) bare-earth elevation (topography) data is available courtesy the U.S. Geological Survey.

@@

\textit{Impervious surface percentage}. This is provided in the National Land Cover Database from the Multi-Resolution Land Charasteristic Consortium \cite{Xian2011-aa}. To generate the impervious surface area (ISA) LandSat data, NLCD land cover, and nighttime light imagery is used. The stable nighttime light intensity is only used to estimate the boundary of of urban areas. The Landsat images were converted to top-of-atmosphere reflectance. The data is provided at a 30m resolution.

\textit{Land cover}. 
Classification of land cover is described in \cite{Homer2015-ce}. 
However, as it is used in the calculation of emissivity when calculating the LST it cannot be used as an independent variable. 
The only landcover that we use is category 11, water.

@@

\textit{Population density}. 
Density allegedly increases the LST \cite{Li2017-yl, Peng2018-cp}, although this result may be the result of confounding with other factors. 
Population density is calculated from the U.S. census at the block level. 
The most recent census was 2010, so we use that data as an estimator of where people reside in the evening. 
This data is also at a lower resolution that the grid cells used, so the grid cell assumes the density of the block that it's centroid is contained within.

\textit{Sky view factor}. Urban canyons have been found to have an effect on UHI because they prevent air circulation \cite{Landsberg1981-mq, Chun2017-mm} and are used to indicate radiation flux within complex environments \cite{Matzarakis2007-xy}. We calculate SVF using \cite{Van_doninck2018-ib} and following the parameters used by \cite{Chun2017-mm}: the number of search directions, $\phi=10^o$; and the radius of the reference circle, $R=300m$. We use a spatial resolution of 6 meters.
% * Yuan 2011, Unger 2004.

\textit{Tree canopy cover}. 
The percent tree canopy cover is calculated using National Agriculture Imagery Program (NAIP) aerial imagery, Landsat 5 imagery, elevation, and existing NLCD data \cite{Coulston2012-uu, Homer2015-ce}. 
The data provided is at a 30m resolution. 
The six reflective bands from Landsat 5 are used to calculate top-of-atmosphere reflectance \cite{Coulston2012-uu} so the data does not contain information used in the LST calculation (which requires radiance). 
% * Rogan 2013, Elmes 2017

\subsection{Data preparation and robustness}
\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{fig/report/holdout_500.png}
    \caption[Holdout cross-validation results at 500 meter resolution]{
    The out-of-bag (OOB) R$^2$ and mean absolute error (MAE) of the models from a 100-fold holdout cross-validation. 
    The models were trained on 80\% of the data and tested on the unseen 20\%.
    When selecting data for the training and testing sets, spatial subsets were used to account for spatial similarities. 
    OOB R$^2$ can vary between $(-\infty, 1)$, where better models have a value near 1. 
    Good models have MAE near 0.
    }
    \label{fig:holdout_500}
\end{figure*}

The objective of our study is to determine the most influential urban characteristics and understand how they relate to land surface temperature.
To do this, we use partial dependence plots to understand how the urban characteristics are associated with land surface temperature.
Given a dataset and for a chosen characteristic, partial dependence is calculated by fixing the value of that characteristic in all observations in the dataset and predicting the land surface temperature. 
This is repeated over a range for the feature of interest.
To evaluate how multiple urban characteristics interact, partial dependence can also be conducted in multiple dimensions.
The swing \cite{Shortridge2015-ub} measures the relative importance of variables by calculating the magnitude of change in temperature associated with each variable.
To calculate partial dependence, we need a data set and a statistical model.

To analyze the data, it needs to transformed into a spatially consistent set.
The raw data is a variety of spatial data types from area level (the population density data), to geostatistical raster (e.g. the land surface temperature).
We choose to resample the data into a grid of square cells.
We conduct the resampling twice to produce a grid of 500m and 100 meter cells. 
Having two grid resolutions allowed us to ensure our conclusions are robust.
For each of the cells the the mean, maximum, and minimum of all variables were calculated. 
To further account for spatial effects of each variable, the mean of the surrounding cells (including diagonal) was calculated and the resulting spatial lag variable was included as an additional independent variable. 
To address multicollinearity between the variables, that would confound our analysis of influence, we remove variables iteratively that have a Variance Inflation Factor greater than approximately 5.
The result is a data set which with we can train a variety of statistical models.

Various statistical models (\S \ref{ss:models}) were trained and then, crucially, validated on unseen data using a technique known as holdout cross-validation. 
This approach partitions 80\% of the data into a training set and the remaining 20\% into the testing set.
The model resulting from the training, is then tested on this unseen test set (known as out-of-bag), to get an estimate of the model accuracy.
This is repeated 100 times and the distribution of the accuracy metric (here we use the mean absolute error (MAE) and variance explained (R$^2$)).
Holdout cross-validation of this type is crucial for statistical analysis to ensure that models are not overfit to data \cite{Geron2017-ek}.
Overfitting occurs when a model fits to the randomness in a dataset, causing it to not be suitable for generalizations. 
When models' accuracy metrics are reported based on in-sample data (the same data it was trained on), the accuracy metrics are high.
This is essential for statistical models and overlooked by the majority of existing studies (e.g. \cite{Zhou2014-wc, Peng2018-cp, Chun2017-mm, Chun2018-so,Wang2019-tree}).
Cross-validation helps to avoid this by evaluating the model on unseen data.

@@

\textit{Random Forest}. A random forest is an ensemble model of \textit{Regression Trees}. 
A regression tree partitions the data based on thresholds for the covariates \cite{Breiman1984-hw}. It continues to divide the data on different covariates until the partitions are as similar as possible. 
The result is a tree-like structure, after which the model is named.
In similar naming fashion, a random forest is a collection of regression trees.
Each tree is \textit{grown} from a random subset of the data and the prediction from the random forest model is the average of the prediction from all of the trees \cite{Breiman2001-rt}.
Tree-based models do not assume linearity and so are generally very flexible and powerful models \cite{Breiman2001-rt, Geron2017-ek}.

@@

\section{Results and Discussion}
In this study we seek to determine the influence and relative importance of urban characteristics on land surface temperature during both the day and night.
To achieve this, the statistical models needed to be accurate.
The results in Figure \ref{fig:holdout_500} indicate that both day and night land surface temperature can be predicted within 1$^o$C at night and 2$^o$C during the day using urban characteristics.
The R$^2$ results, also based on unseen data, suggest that more than 80\% of the data variance is captured by the models for the mean LST during both night and day, while the mean \% variance explained is generally between 50 and 80\% for the maximum temperatures.
Compared to models presented in similar studies, these are strong results, especially given that the existing studies failed to validate their models on unseen data.
We find that the most accurate model is the random forest, closely followed by the gradient boosted regression trees.
The weakest model is the linear regression, incidentally the one that is used in the majority of existing studies into land surface temperature.
This strong result allows us to assess the influence and relative importance that these characteristics.

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{fig/report/pdp_500.png}
    \caption[Partial dependence plots for LST at 500 meter resolution]{
    Partial dependence plots show how the land surface temperature ($^oC$, y axis) changes with each urban characteristic as the other variables are held at their average (mean) value. 
    The left hand side shows the effect each variable has on the (a) mean land surface temperature (LST) during the night, (b) maximum LST during the night, (c) mean LST during the day, (d) maximum LST during the day. 
    Each of the models are shown and this indicates the model uncertainty in the relationships.
    There are multiple lines for each model based on bootstrap samples of the data, which indicates the data uncertainty.
    The histograms on the $x$-axis shown the distribution of the observed data.
    This is for the 500 meter resolution.
    }
    \label{fig:pdp_500}
\end{figure*}

Relative variable importance for all of the urban characteristics, calculated with \textit{swing}, is shown in Figure \ref{fig:importance_500}. 
The characteristics, shown on the $y$ axis are ordered by the mean swing across all of the models, and the models are ordered left-to-right by their cross-validated accuracy.
The most important characteristics during the night are tree canopy cover, greeness (nbdi), water, the sky view factor, and the elevation (digital surface model).
This is consistent for both the mean and maximum temperature during the night.
It is also generally consistent between the models. 
In contrast, during the day, albedo (surface whiteness) is more important and the sky view factor is less influential.
Albedo's small association with nighttime temperatures is surprising given that darker surfaces store heat during the day that is released during the night \cite{Voogt2003-mm, Zhou2014-wc}.
This, and the other influences can be examined using partial dependence.

\textit{Vegetation and impervious surfaces}. 
During the day, increasing impervious surfaces and decreasing vegetation causes increased sensible heat flux and lowered latent heat flux \cite{Voogt2003-mm, Peng2012-iy, Zhou2014-wc}.
This is thought to be less important during the night because latent and sensible heat dominant during the day, but ground heat flux dominates at night \cite{Zhou2014-wc, Voogt2003-mm}.
Additionally, vegetation is expected to reduce daytime temperature due to transpiration which increases the latent heat flux \cite{Zhou2014-wc}. 
However, because transpiration does not occur at night, vegetation's effect on night temperatures is debated. 
In our study, percentage tree canopy and impervious surface cover must be discussed together because they are 100\% correlated (Figure \ref{fig:imp_tree}).
To distinguish between vegetated surfaces and other pervious surfaces, we also include NDVI, a measure of a surface's greeness.
At night, we find that \% tree canopy is the most influencial characteristic on land surface temperature, followed by NDVI (Figure \ref{fig:pdp_500}).
These partial dependence results indicate a decrease in up to 7$^oC$ as the percentage area of trees or other pervious surface increases.
This reduction in temperature could be due to a reduction in the impervious surfaces that store heat.
However, we also observe that as the NDVI increases, the land surface temperature decreases.
The two-dimension partial dependence (Figure \ref{fig:pdp_2dnight_500}) evaluates how LST varies with both characteristics.
Figure \ref{fig:pdp_2dnight_500}d shows how the LST changes as both tree canopy/impervious surface and the NDVI change.
The greener and more pervious a surface is, the cooler it is during the night. 
The total change here is also approximately 7$^oC$, which is a considerable amount given these two-dimensional partial dependence are calculated using the random forest model, which is the less variable of the five models shown in Figure \ref{fig:pdp_500}.
These results contradict standing conclusions that vegetation has no effect on night time LST \cite{Peng2012-iy, Zhou2014-wc};
rather, vegetation appears to be strongly associated with lower land surface temperatures during the night.

During the day, we again see that vegetation and impervious surfaces are associated with the greatest reduction in land surface temperature, with approximately 10$^oC$ change (Figure \ref{fig:pdp_2dday_500}d). 
This daytime result is consistent with expectation \cite{Chun2017-mm, Peng2018-cp, Wang2019-tree,Zhou2014-wc, Chun2018-so}.
Because not all pervious surface types have the same cooling potential, and because of trade-offs with irrigation of vegetated surfaces \cite{Gober2009-im}, understanding the ideal vegetated surface type for day and night temperatures is a future step.
